---
title: "ISTA 320 Final Project"
author: "Edoardo Martinelli"
date: "10/13/2021"
output: html_document
---

For my final project I decided to make three visualizations of data breaches from 2004 to 2021.
The data set includes variables which are Entity (the name of the company), Year, Records (Records lost from data breach),
Organization.Type (What type of company), and Method (What method was used for data breach). 

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
library(scales)

data_breach <- read.csv("data/DataBreaches(2004-2021).csv")
```

For my first graph I decided to do a line plot of total loss in records per year.
We can see that 

```{r}
# First
data_breach %>%
  group_by(Year) %>%
  summarize(tot_lost = sum(Records)) %>%
  filter(Year != "2014 and 2015", Year != "2018-2019") %>% 
  ggplot(aes(x = Year,
             y = tot_lost,
             group = 1)) +
  geom_line() +
  scale_y_continuous(labels = label_number(suffix = " B", scale = 1e-9)) +
  ylab("Total Loss in Records") +
  labs(title = "Plot of total loss in records per year")
```


```{r}
#2
data_breach %>% 
  group_by(Organization.type) %>% 
  summarize(tot_lost = sum(Records)) %>%
  slice_max(order_by = tot_lost, n = 10) %>%
  ggplot(aes(fct_rev(fct_reorder(Organization.type, tot_lost)),
             tot_lost,
             fill = Organization.type)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = label_number(suffix = " B", scale = 1e-9)) +
  xlab("Organization Type") +
  ylab("Total Loss in Records") +
  labs(title = "Plot of total loss in records by Organization type",
       fill = "Organization Type")
```

```{r}
#3
method_percentage <- data_breach %>%
  filter(!is.na(Method)) %>%
  group_by(Method) %>%
  summarize(n = n()) %>%
  mutate(total = sum(n),
         percentage = n/total,
         Method = factor(Method))

method_percentage %>%
  slice_max(order_by = percentage, n = 5) %>%
  ggplot(aes(fct_rev(fct_reorder(Method, percentage)),
             percentage,
             fill = Method)) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Organization Type") +
  ylab("Percentage") +
  labs(title = "Plot of percentage of which data breach methods used",
       fill = "Data Breach Method")
```

